The stopping power of \clau{hydrogen in} hafnium and the importance of relativistic $4f$ electrons

### 2020-04-20 response.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\setlength{\parindent}{0pt}
\usepackage{soul}
\usepackage{xcolor}
\usepackage[margin=0.8in]{geometry}
\usepackage{graphicx}

\setstcolor{red}

% Title Page
\title{Response to referee reports \\ AP11756/Montanari }
%\author{C. C. Montanari, A. M. P. Mendez, D. M. Mitnik, J. E. Miraglia, \\ P.A. Miranda, R. Correa, J. Wachter, M. Aguilera, \\ E. Alves, N. Catarino and R.C. da Silva}
%\date{\today}

@@

\begin{document}
\maketitle

% ----------------------------------------------------------------------
\section{Reply to report of the First Referee}
% ----------------------------------------------------------------------

\textsl{1. Since only protons are considered as projectiles, the title 
needs to be modified accordingly.}

\vspace{0.1cm}
{\color{red} The title has been modified following the referee's recommendation:}
{\small ''The stopping power of hydrgen in hafnium and the importance of relativistic $4f$ electrons''}.

% ----------------------------------------------------------------------
\vspace{0.35cm}
\textsl{2. It appears that calculations assume bare protons. According 
to CasP, neutrals play a considerable role in charge equilibrium around 
and below the Bragg peak. Some explanation is needed here.}

\vspace{0.1cm}
{\color{red} It is correct, our calculations assume bare protons. We consider that there is no bound electron or very small probability because free electrons of metals screens the hydrogen and this screening is enough at low energies as to have no bound electron. CasP includes a very good fitting of experimental charge states and a powerful scaling rule about them. However, it is based on measurements of charge states outside the metal. For multi-charged ions almost no differences are expected and CasP charge state works very well. For H and He ions, differences are important.  We add the following explanation about why considering just proton impact, and the reference to CasP and the paper of charge states by Schwietz and Grande.}

\vspace{0.1cm}
{\small ''Finally, in all our calculations we assumed the projectile to be proton and not neutral hydrogen. When an ion moves inside a metal, the FEG screens the nucleus, so the binding energies will be smaller than outside the metal, and this effect is more important a low impact velocities $v$. In the case of hydrogen the difference is drastic, i.e. for H inside Hf ($rs=2.07$), the $1s$-bound state is almost null at $v<2$. It is worth to mention that this assumption agrees with Ziegler SRIM code, but differs from CasP code that predicts neutral hydrogen at very low velocities.''}
% ----------------------------------------------------------------------

\vspace{0.35cm}
\textsl{3. Thickness measurements by Rutherford backscattering rely 
‘heavily’ on the scattering cross section, while energy loss is a minor 
correction depending on foil thickness (p.1, 2nd paragraph).}

\vspace{0.1cm}
{\color{red} Pedro Miranda ..........................................................}
\vspace{0.2cm}

@@

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{4. In order to get a more precise thickness value of the Hf 
foil, the authors use the energy loss of 5.486 MeV alphas as a standard, 
with SRIM as a reference. SRIM is based on empirical data, and for alpha 
particles in Hf, the IAEA database lists one single data point for He 
in Hf. Considering the spread in published stopping cross sections for 
e.g. He in Au, listed in the IAEA database, some explanation is needed 
to justify an error estimate of only 5\% (1st paragraph in section 
IIB).}

\vspace{0.1cm}
{\color{red} The referee is correct, we have used the stopping of 5.486 MeV alphas and SRIM as reference to get the thickness. Although SRIM fits the experimental data available, it is not just a fitting, it is a mix of model and fitting (semiempirical). At high impact energies where the stopping is well known srim is quite reliable. Figure \ref{HeHf} here can clarify these assertions. 
a) In this figure we plotted together H and He in Hf, dividing by $Z^2$. Above 1 MeV/amu both srim curves and our theoretical results agree within $2\%$. The vertical pink line indicates the impact energy we use as reference. It is in the region where the srim predictions are reliable. b) srim is semiempirical, the comparison of H and He in the same target shows clearly how srim for H ions curves down and up to fit Sirotonin data. It is also clear how srim for He ion overvalue in order to fit the single measurement by Chu, Ziegler and others. Why do we say "overvalue"? Because the stopping of He ions divided by 4 can be equal or below the stopping of H ions. }

%-----------------------------------------------------------------------
\begin{figure}[!t]
\centering
\includegraphics[width=12.cm]{Fig_HeHf.eps}
\caption{Comparison of stopping of Hf for H and He at high impact energies}
\label{HeHf}
\end{figure}
%-----------------------------------------------------------------------

% ----------------------------------------------------------------------
\vspace{0.35cm}
\textsl{5. In connection with the discussion of relativity and outer-shell electrons (p.3) it might be instructive to mention that 
this is the reason for the existence of relativistic quantum chemistry programs. Since the authors find it surprising that the effect increases in magnitude from inner to outer shells, the reader needs an explanation. I should think the effect of deviations in the screening add up shell by shell, but I may be wrong. Moreover, the sign of the discrepancy in figure 2a deserves attention, which seems to change from positive up to 4p- to negative from 5p+.}

\vspace{0.1cm}
{\color{red} The referee is correct. We included the following modifications in the text
(page 3, lines 210-228):}

\vspace{0.1cm}
{\small ``In the comparison with the experimental values in Fig. 2 (a) it can be noted that the sign of the binding energy deviations is inverted for the outer $5s$-$4f\pm$ electrons, with the experimental binding energies being less bounded than our theoretical ones. Small differences for the outer shells are expected since  the experimental
values correspond to hafnium in solid-state, while our theoretical  calculations correspond to the element in the gas phase.
For more detail, Fig. 2 (b) shows  the relative errors
with respect to the experimental  values.
This figure shows clearly that the relativistic corrections are very important to describe
the atomic structure of hafnium, even for the outer shells.
It turns out that the errors committed in the non-relativistic calculations of the inner shell orbitals 
propagate, through the Hartree-Fock approximation, to the outer shells.The importance of full-relativistic calculations for the outer shells has already been noted before for Au, Pb and Bi and W.''}

%{\small ``One would expect the relativistic contributions to only affect  significantly the inner shells. The relativistic binding energies of the  electrons belonging to these shells are 2.5\% from the experimental values, while the non-relativistic ones are within $\sim$10\%. Moreover, we notice the relativistic corrections greatly influences the outer $4f$ shell. This effect may be a consequence of the orthogonality with the  relativistic inner shells. The relativistic $1s$-$4d$ electrons are little but more bounded than the experimental values. This sign is inverted for the outer electrons, being less bounded experimentally than the relativistic calculations.  This change of sign can be attributed to the fact that the experimental values correspond to hafnium in solid-state, while our theoretical calculations correspond to the element in the gas phase.''}

% ----------------------------------------------------------------------
\vspace{0.35cm}
\textsl{6. The importance of 4f-5p screening is documented clearly (p.3), but the reader will ask why this does not affect other shells. I assume that the effect is important when two or more subshells have similar energies and, hence, overlap in real and velocity space. If so, I suggest to mention this. If not, another explanation is needed.}

\vspace{0.1cm}
{\color{red} The mixing criteria is explained on page 3: we consider the quantum uncertainty in energy so bound electrons with similar energies and mean radio $<r>$ response as a single density of electrons including screening among them. We recognize that a more detailed study is needed, we are developing such work as a systematic for different targets involving 4f electrons. What we have done here for Hf is to check the importance of the contribution to the total stopping. Only the 5p-4f mixing have influence in the total stopping because they are the second most important contribution after the FEG. Some other shells may be mixed, and we calculate them, but the total stopping changes in very small $\%$.  We added lines 262-264 to page 4.}

{\small '' As already mentioned, at higher energies inter-shell screening is possible for other subshells (i.e. 4p-4d for impact energies above 0.9 MeV), but its weight in the total stopping is minor for deeper shells.'' }

% ----------------------------------------------------------------------
\vspace{0.35cm}
\textsl{7. In view of the large contribution of the outermost shell, I 
suggest to include the core contribution in figure 3.}

\vspace{0.1cm}
{\color{red} We thank the referee for this suggestion, actually we had this figure, but we doubted on including all the curves. Now Fig. 3 includes separate FEG and bound electron contribution, and the total curves too.}

% ----------------------------------------------------------------------
\vspace{0.35cm}
\textsl{8. In addition to comparisons with experiments and empirical 
models, it would be instructive to see comparisons with theoretical 
models like CasP and DPASS in figure 4.}

\vspace{0.1cm}
{\color{red} We thank this suggestion. We had already compared with Casp5.2 and DPASS for ourselves before submitting. In this version we added CasP5.2 and DPASS to figure 4, we added a mention in the introduction(last paragraph, lines 68-73), and the following comment in page 6, lines 315-322:}

{\small ''The stopping maximum is a very sensitive region for any full theoretical description, with no parameters at all, and this is quite visible in a linear-scale plot like Fig. [4]. However, the impact energy for the maximum seem to agree between our curve and DPASS, though $10 \%$ below. Instead, CasP predicts a stopping maximum $10 \%$ above and a completely different shape at lower energies.''}

@@

\newpage
% ----------------------------------------------------------------------
\section{Reply to report of the Second Referee}
% ----------------------------------------------------------------------

\textsl{1. Can the authors discuss how possible impurities in the sample and -especially- on both sides of the target would affect the final
stopping data? The authors must discuss it in the manuscript.}

\vspace{0.1cm}
{\color{red} Pedro Miranda, ..........................................................}
\vspace{0.2cm}

@@

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{3. What is accuracy of the primary beam energy of the accelerator?
Moreover: when the target is placed and removed from the front of the
detector, does the detector itself need to unbiased and the chamber
opened? If so, might there be any offset ($\sim$keV) in the detection
system that would affect the energy distribution peaks, hence the
Gaussian position from the fits (Fig. 1)?}

\vspace{0.1cm}
{\color{red} Pedro Miranda, ..........................................................}
\vspace{0.2cm}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{4. The experimental data above $\sim$1.5 MeV is systematically 
below the data from Ref. [7], hence SRIM semi-empirical approach, and 
also below the authors’ theory output. At these energies (and higher), 
the agreement between experiment and theory is expected to be better than
a few percent (that is why you used stopping power for 5.486 MeV alpha
on Hf to obtain the target thickness). Can the authors explain why
there is such systematic difference for the present measurements using
protons?}

{\color{red} There is a small systematic difference between our experiments and theory for impact energies above 1.5 MeV, but it is within the experimental error. Figure~\ref{highE} here, amplifies the results shown in the manuscript by focusing  on the high energy region. Note that we included our theoretical results, Srim, Casp5.2 and DPASS to this figure. The difference is actually small, being less than $3\%$ for the two values above 2 MeV. We included the following sentence in the text........................Pedro, PLEASE CHECK THIS}

%-----------------------------------------------------------------------
\begin{figure}[!t]
\centering
\includegraphics[width=12.cm]{Fig04.eps}
\caption{stopping at high impact energies, theoretical curves and symbols for data explained inside the figure}
\label{highE}
\end{figure}
%-----------------------------------------------------------------------

@@

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{5. The authors claim an overall uncertainty of 5\%, but it is not clear
how they obtained this value, or whether only statistical
contributions were taken into account. Nevertheless, uncertainties of
measurements have to be treated following the Guide to the Expression
of Uncertainty in Measurement, International Organization for
Standardization, Geneva Switzerland, 1995. Especially for stopping
power measurements in transmission approach, there are other sources
of uncertainties that might potentially affect the final results
systematically (see comments 1, 2 and 3), and they must be discussed
in more details.}

\vspace{0.1cm}
{\color{red} Pedro Miranda, ..........................................................}
\vspace{0.2cm}

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{6. Can the present theory in the manuscript be compared to other
models (e.g., CasP and DPASS), especially towards low energies? Aiming
thus a constructive discussion on the overall-agreement among
different models to predict SCS of “challenge” materials, as
transition metals (f-subshell)?}

\vspace{0.1cm}
{\color{red} We thank the referee for this suggestion that improves this new version.  We have added CasP5.2 and DPASS curves in Figure 4, we have also added a mention in the introduction (last paragraph, lines 68-73), and the following comment in page 6, lines 315-322:}

{\small ''The stopping maximum is a very sensitive region for any full theoretical description, with no parameters at all, and this is quite visible in a linear-scale plot like Fig. [4]. However, the impact energy for the maximum seem to agree between our curve and DPASS, though $10 \%$ below. Instead, CasP predicts a stopping maximum $10 \%$ above and a completely different shape at lower energies.''}

@@

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{7. In the last sentence of Sec. IV, the authors point out for a need
of more experimental data below 100 keV H+ but, in my opinion, more
data are needed even at higher energies, e.g., nearby (and slightly
above) the Bragg peak... Please enhance discussions about it.}

\vspace{0.1cm}
{\color{red} We agree with this comment by the referee. In a ion-target system with only two set of measurements more data is needed. In the energy region 80-500 keV only the data by Sirotinin group is available and cannot be compared with others. SRIM is not a test for this data, SRIM includes this data in the adjust (more detail about this can be found in the answer 4 to the first referee report). We change the comments in page 6 lines 328-336 and lines 364-366 as follows:}

{\small ''Future experiments  would be important to
complete this study on the stopping of Hf for protons. In the energy region $80-500$ keV only the measurements by Sirotinin group in 1984 are available. SRIM clearly includes these values in the adjust of the code, so it does not represent a test for this data. Undoubtedly, new measurements are needed in three specific energy regions: below 500 keV, around the stopping maximum (i.e. $50-200$ keV) and at low impact energies''}

{\small ''Future experiments for impact energies 
below 500 keV, around the stopping maximum and in the low energy region would be important to complete this study.''}

@@

\newpage
% ----------------------------------------------------------------------
\section{Reply to report of the Third Referee}
% ----------------------------------------------------------------------

\textsl{
- Regarding the theoretical model that describes the contribution to
the stopping power of the bound electrons, the authors emphasize the
importance of including relativistic effects in the description of the
binding energies and densities of the Hf 1s-4f states. However, it is
not clear if the calculations performed by the authors are done for
gas phase Hf atoms or for the solid Hf as in the experiments of Ref.
[38]. In case of being for the former, the slightly worse agreement
found for the 4f states as compared to the experimental values could
be related to this point.}

\vspace{0.1cm}
{\color{red} The referee is correct, our calculations are performed for gas phase Hf atoms. We have added the following paragraph in this version of the manuscript (page 3, lines 207-215) to express this clearly.}

{\small ''In the comparison with the experimental values in Fig.2 (a) it can be noted that the sign of the binding energy deviations is inverted for the outer $5s$-$4f\pm$ electrons, with the experimental binding energies being less bounded than our theoretical ones. Small differences for the outer shells are expected since  
the experimental values correspond to hafnium in solid-state, while our theoretical  calculations correspond to the element in the gas phase.''}

@@

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{
1. The results from the three different theoretical calculations shown
in Fig. 3 are not completely clear. The authors emphasize the
importance of considering both screening effects and relativistic
effects for the 4f and 5p electrons, but it is not clear if later in
the text the authors use indistinctly the terminology ``relativistic''
and ``screening'' to refer to the same issue. For instance, are the ML
results shown in Fig. 3 calculated with the relativistic binding
energies and densities, but with and without including screening
corrections, or do the two ML curves differ in whether the
relativistic corrections are or not included? It is neither clear what
description of the bound electrons is used in the curve labeled as
``SPCC(FEG)+SLPA(bound)''.}

\vspace{0.1cm}
{\color{red} Thank you for the comment that helps us to clarify this point. Our SLPA calculations use the relativistic wave functions and binding energies for the bound electron contribution to the stopping power.The three curves in Fig. 3 and the curve in Fig. 4 are relativistic in this sense. We have added an explicit mention about the use of relativistic wave functions and binding energies in page 4 lines 200-201, and also this comment in page 4, lines 287-289:}

{\small ''Bound $1s$-$4f$ electrons (relativistic wave functions and binding energies) are calculated with the SLPA and shown separately in Fig. 3''}

@@

% ----------------------------------------------------------------------
\vspace{0.25cm}
\textsl{
2. If all the results of Fig. 3 were calculated for the relativistic
description, it would be also interesting to show the results obtained
with the non-relativistic description in order to demonstrate the
importance of this correction.}

\vspace{0.1cm}
{\color{red}  The SLPA calculations for the bound electrons uses the Hartree-Fock wave funtions and binding energies, which are available in the tables by Clementi-Roetti or by Bunge for elements up to Xe (Z=54). The errors in the non-relativistic calculations worsen for higher Z atoms, and this is the reason why these tables do not extend any further than Z=54.  The fail of non-relativistic values of binding energies is shown in Fig. 2. We have added a comment on this in page 3, lines 214-228. }

\vspace{0.1cm}
{\small ``For more detail, Fig. 2 (b) shows  the relative errors
with respect to the experimental  values.
This figure shows clearly that the relativistic corrections are very important to describe
the atomic structure of hafnium, even for the outer shells.
It turns out that the errors committed in the non-relativistic calculations of the inner shell orbitals 
propagate, through the Hartree-Fock approximation, to the outer shells.The importance of full-relativistic calculations for the outer shells has already been noted before for Au, Pb and Bi and W.''}

{\color{red} When dealing with stopping in targets with Z>54 we need to face our own calculation of the wave functions and binding energies. And we do it within the relativistic approximation to the best of our possibilities. However, one can wonder if such discrepancy in the non-relativistic results affects the stopping power. First, it will only affect the bound electron contribution to the total stopping. Second, at high energies (i.e. 1 MeV) the results converge to Bethe limit for any density of electrons as far as the correct number of target electrons is included. Having clarified this, we include in the figure \ref{NR} here, the total stopping using the non-relativistic values we calculate using the autostructure code. As can be noted, at sufficiently high energies it is the same, but differs around the stopping and at lower impact energies. We have included in this figure and in Fig. 4 of the manuscript the DPASS results by Sigmund.}

### 2020-04-20 Stopping_Hf_abril20.tex
checked with grammarly

\documentclass[aps,pra,reprint,groupedaddress,showpacs,showkeys]{revtex4-1}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath} 
\usepackage{graphicx}
\usepackage{soul}
\usepackage{xcolor}
\setstcolor{red}
\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

%\def\clau#1{\textcolor{ForestGreen}{#1}}
%\def\minor#1{\textcolor{green}{#1}}

@@

%Title of paper
\title{Stopping power of hafnium by proton impact and \\
the importance of relativistic $4f$ electrons}
%Authors and affiliations

%---------------------------Argentina
\author{C. C. Montanari}
\email{mclaudia@iafe.uba.ar}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
Argentina}
%------------------------------
\author{A. M. P. Mendez}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
Argentina}
%-------------------------------
\author{D. M. Mitnik}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
Argentina}
%----------------------------------------------
\author{J. E. Miraglia}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
Argentina}
%---------------------------Chile

\author{P.A. Miranda}
\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
7800002, Chile}
%----------------------------
\author{R. Correa}
\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
7800002, Chile}
%----------------------------
\author{J. Wachter}
\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
7800002, Chile}
%----------------------------
\author{M. Aguilera}
\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
7800002, Chile}
%-------------------------Lisboa
\author{E. Alves}
\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
Portugal}
\affiliation{Instituto de Plasma e Fus\~{a}o Nuclear, Instituto Superior 
T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, Portugal}
%---------------------------
\author{N. Catarino}
\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
Portugal}
%---------------------------
\author{R.C. da Silva}
\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
Portugal}

@@

So far, only one experimental work has been published regarding the 
stopping power cross section of pure hafnium for protons~\cite{Sirotinin}, 
while more attention has been recently given to studies involving 
hafnium oxide due to its practical use~\cite{Abril,Behar,Primetzhofer,Roth}. 
It is well known that significant attention has been paid in recent 
years to transition metal-oxides such as HfO$_2$ because of their 
potential as alternative gate dielectrics to replace SiO$_2$ for the 
future generation of nano-electronics with less than 45 nm gate 
length~\cite{Choi,Robertson}. Some important physical properties of the 
above mentioned metal-oxide films depend on its thickness, which is 
often measured by using Rutherford Backscattering Spectrometry~\cite{Alfassi01,Tesmer01}, 
a method that relies heavily on the determination of the stopping power 
of ion beams in the material of interest.

@@

The theoretical approach implemented in this work uses the shell-wise 
local plasma approximation (SLPA)~\cite{mon13} to describe the energy 
transferred to the bound $1s$-$4f$ electrons, and two different models 
for the FEG: the screened potential with cusp condition model (SPCC)~\cite{mon17}, 
which is non-linear binary formalism; and the Mermin-Lindhard dielectric 
formalism (ML)~\cite{Mermin}, for energies around the stopping maximum 
and above. Our model requires the relativistic wave functions and 
binding energies of Hf, and considers 4 electrons per atom in the 
FEG~\cite{mendez2019}. Hf has the extra interest of the filled 
$4f$-subshell (with 14 electrons) as the main contribution below the 
FEG, causing the stopping cross sections to be very sensitive to a good 
description of this shell. The screening among the $4f$ and $5p$ 
electrons has been considered and found to play a major role within the 
SLPA calculations.

The experimental details and data are given in Section~\ref{experiment}, 
while the theoretical method is explained in Section~\ref{theory}. 
Present theoretical and experimental values are finally compared to the 
available experimental values~\cite{Sirotinin}, the ICRU-49 
calculations~\cite{ICRU49} and the SRIM-2013 package~\cite{Ziegler01}. 
Conclusions and discussions are given in Section~\ref{conclusion}. All 
the present data can be found at the Zenodo platform \cite{zenodo}.

@@

%=======================================================================
\section{Theoretical Method} 
\label{theory}
The energy loss of ions in metal targets responds to different physical 
mechanisms, depending on the impact ion velocity. At low velocities, the 
binary collisions are responsible for the loss of energy by the ion. 
The main contribution is the ionization of electrons of the metal 
conduction band, which is well approximated by a free electron gas (FEG) 
of Fermi velocity $v_F$. Above certain velocity (i.e. $v\geq 1.5\,v_F$), 
not only binary but also collective excitations (plasmons) 
occur~\cite{mon17}. Moreover, at high energies, also the bound electrons 
contribute to the stopping power. The method used in this work combines 
a description for the interaction with the valence (or conduction) 
electrons as a FEG, and a different one for the interaction with the 
bound electrons.

@@

Hafnium ($Z=72$, [Xe] $4f^{14}\,6s^2\,5d_{3/2}^1\,5d_{5/2}^1$), belongs 
to the first groups of transition metals, with four electrons as FEG 
(theoretical $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound. We compared 
the computed $r_S$ with the experimental value obtained from measured 
energy loss function by Lynch \textit{et al} \cite{lynch75}. The 
experimental plasmon energy of Hf is $\hbar\omega_P \approx 15.8$ eV, 
with a width at half maximum $\delta \approx 4.4$ eV, and 
$r_S \approx 2.07$ a.u. \cite{lynch75}. The difference of less than 
$5\%$ between theoretical and experimental $r_S$ assess Hf as a 
canonical target~\cite{mon17}.

@@

%----------------------------------------------------------------------
\begin{figure*}[!t]
\centering
\includegraphics[width=11.cm]{bindener.eps}
\caption{(a) Binding energies of Hf. Non-relativistic and relativistic 
calculations are given with filled symbols. Experimental measurements 
for solids~\cite{williams1995} are depicted with open circles. 
(b) Corresponding relative errors respect to experimental data.}
\label{Binding_E}
\end{figure*}
%------------------------------------------------------------------------

For the stopping power due to bound electrons, the SLPA~\cite{mon17,mon13} 
is employed. It is worth mentioning that the only inputs for the SLPA 
are the space-dependent densities of each shell in the ground state, 
and their binding energies. Collective processes and screening among 
electrons are included. Since hafnium is a relativistic target, the 
wave functions and binding energies must be obtained by solving the 
many-electron Dirac Hamiltonian. Details of these calculations have been 
published in \cite{mendez2019}, while Slater-type orbital expansions are 
given in \cite{Hf_arxiv}.

>>>>

\documentclass[aps,pra,reprint,superscriptaddress]{revtex4-1}
%\documentclass[aps,pra,reprint,groupedaddress,showpacs,showkeys]{revtex4-1}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath} 
\usepackage{graphicx}
\usepackage{soul}
\usepackage{xcolor}
\setstcolor{red}
\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

\def\clau#1{\textcolor{red}{#1}}
\def\duda#1{\textcolor{green}{#1}}
\def\ale#1{\textcolor{blue}{#1}}

@@

%Title of paper
\title{The stopping power of \clau{hydrogen in} hafnium and the importance of relativistic $4f$ electrons}
%Authors and affiliations

%---------------------------Argentina
\author{C. C. Montanari}
\email{mclaudia@iafe.uba.ar}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
Argentina}

%---------------------------Chile
\author{\clau{P. A. Miranda}}
\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
7800002, Chile}
%-------------------------Lisboa
\author{\clau{E. Alves}}
\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
Portugal}
\affiliation{Instituto de Plasma e Fus\~{a}o Nuclear, Instituto Superior 
T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, Portugal}

%------------------------------Argentina
\author{A. M. P. Mendez}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
Argentina}
%-------------------------------
\author{D. M. Mitnik}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
Argentina}
%----------------------------------------------
\author{J. E. Miraglia}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
Consejo Nacional de Investigaciones Cient\'{\i}ficas y T\'{e}cnicas - 
Universidad de Buenos Aires, Pabell\'on IAFE, 1428 Buenos Aires, 
Argentina}

%---------------------------- Chile
\author{R. Correa}
\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
7800002, Chile}
%----------------------------
\author{J. Wachter}
\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
7800002, Chile}
%----------------------------
\author{M. Aguilera}
\affiliation{Departamento de F\'isica, Facultad de Ciencias Naturales, 
Matem\'atica y del Medio Ambiente. Universidad Tecnol\'ogica Metropolitana. 
7800002, Chile}

%-------------------------Lisboa
\author{N. Catarino}
\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
Portugal}
%---------------------------
\author{R. C. da Silva}
\affiliation{Centro de Ci\^{e}ncias e Tecnologias Nucleares, Instituto 
Superior T\'ecnico, Universidade de Lisboa, 2696-953 Sacav\'{e}m, 
Portugal}

@@

So far, only one experimental work has been published regarding the 
stopping power cross section of pure hafnium for protons~\cite{Sirotinin}, 
while more attention has been recently given to  studies involving 
hafnium oxide due to its practical use~\ale{\cite{Abril,Behar,Primetzhofer,Roth}}. 
It is well known that significant attention has been paid in recent 
years to transition metal-oxides such as HfO$_2$ because of their 
potential as alternative gate dielectrics to replace SiO$_2$ for the 
future generation of nano-electronics with less than 45 nm gate 
length~\cite{Choi,Robertson}. Some important physical properties of the 
above mentioned metal-oxide films depend on its thickness, which is 
often measured by using Rutherford Backscattering Spectrometry~\cite{Alfassi01,Tesmer01}, 
a method that relies heavily on the determination of the stopping power 
of ion beams in the material of interest.

@@

The theoretical approach implemented in this work uses the shell-wise local plasma approximation (SLPA)~\cite{mon13} to describe the energy 
transferred to the bound $1s$-$4f$ electrons, and two different models 
for the FEG: \clau{in low energy region,}
the screened potential with cusp condition model (SPCC)~\cite{mon17}, 
which is non-linear binary formalism; and the Mermin-Lindhard dielectric 
formalism (ML)~\cite{Mermin}, for energies around the stopping maximum 
and above. Our model requires the relativistic wave functions and 
binding energies of Hf, and considers 4 electrons per atom in the 
FEG~\cite{mendez2019}. Hf has the extra interest of the filled 
$4f$-subshell (with 14 electrons) as the main contribution below the 
FEG, causing the stopping cross sections to be very sensitive to a good 
description of this shell. The screening among the $4f$ and $5p$ 
electrons has been considered and found to play a major role within the 
SLPA calculations.

The experimental details and data are given in Section~\ref{experiment}, 
while the theoretical method is explained in Section~\ref{theory}. 
Present theoretical and experimental values are finally compared to the \clau{only experimental values by Sirotinin group~\cite{Sirotinin} measured with backscattering method, the theoretical results by  Grande and Schiwietz~\cite{Grande,casp52}, and by Sigmund and Schinner~\cite{DPASS20}, and also with the semi-empirical values from the SRIM-2013 package~\cite{Ziegler01} and the ICRU-49 tabulation~\cite{ICRU49}}. 
Conclusions and discussions are given in Section~\ref{conclusion}. All 
the present data can be found at the Zenodo platform \cite{zenodo}.

@@

%=======================================================================
\section{Theoretical Method} 
\label{theory}
The energy loss of ions in metal targets responds to different physical 
mechanisms, depending on the impact ion velocity. At low velocities, the 
binary collisions are responsible for the loss of energy by the ion. 
The main contribution is the ionization of electrons of the metal 
conduction band, which is well approximated by a free electron gas (FEG) 
of Fermi velocity $v_F$. Above certain velocity (i.e. $v\geq 1.5\,v_F$), 
not only binary but also collective excitations (plasmons) 
\ale{occur}~\cite{mon17}. Moreover, at high energies, also the bound electrons 
contribute to the stopping power. The method used in this work combines 
a description for the interaction with the valence (or conduction) 
electrons as a FEG, and a different one for the interaction with the 
bound electrons.

@@

Hafnium ($Z=72$, [Xe] $4f^{14}\,6s^2\,5d_{3/2}^1\,5d_{5/2}^1$), belongs 
to the first groups of transition metals, with \ale{four} electrons as FEG 
(\st{theoretical} $r_S=2.14$ a.u.) and $1s$-$4f$ electrons bound. We compared 
the computed $r_S$ with the experimental value obtained from measured 
energy loss function by Lynch \textit{et al} \cite{lynch75}. The 
experimental plasmon energy of Hf is $\hbar\omega_P \approx 15.8$ eV, 
with a width at half maximum $\delta \approx 4.4$ eV, and 
$r_S \approx 2.07$ a.u. \cite{lynch75}. The difference of less than 
$5\%$ between theoretical and experimental $r_S$ assess Hf as a 
canonical target~\cite{mon17}.

@@

%----------------------------------------------------------------------
\begin{figure*}[!t]
\centering
\includegraphics[width=11.cm]{bindener.eps}
\caption{(a) Binding energies of Hf. \clau{Present relativistic and available non-relativistic~\cite{badnell97}  
values} are given with filled symbols. Experimental measurements 
for solids~\cite{williams1995} are depicted with open circles. 
(b) Corresponding relative errors respect to experimental data.}
\label{Binding_E}
\end{figure*}
%------------------------------------------------------------------------

For the stopping power due to bound electrons, the SLPA~\cite{mon17,mon13} 
is employed. It is worth \ale{mentioning} that the only inputs for the SLPA 
are the space-dependent densities of each shell in the ground state, 
and their binding energies. Collective processes and screening among 
electrons are included. \ale{Since} hafnium is a relativistic target, the 
wave functions and binding energies must be obtained by solving the 
\ale{many-electron} Dirac Hamiltonian. Details of these calculations 
\clau{and a table of binding energies} have been published in 
Ref.~\cite{mendez2019}, while Slater-type orbital expansions are 
given in \cite{Hf_arxiv}.
